---
title: "Assignment 4"
author: "Madeline Berger"
date: "11/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load necessary packages 

library(tidyverse)
library(vcdExtra)
library(kableExtra)
library(RColorBrewer)
library(car)
library(effsize)
library(kableExtra)
```

Read in the CSVs

```{r}


lobster_size <- read_csv("lobster_size_abundance.csv")
lobster_traps <- read_csv("lobster_traps.csv")
```

The lobster_size file needs to be converted to tidyverse format

```{r}

#convert data to tidyverse
lobster_size1 <- as.data.frame(lobster_size)
lobster_size2 <- expand.dft(lobster_size1, freq = "COUNT") %>% 
  select(YEAR, SITE, SIZE)

lobster_traps1 <- as.data.frame(lobster_traps)
lobster_traps2 <- expand.dft(lobster_traps, freq = "TRAPS")
```

```{r}

#Data wrangling into tables and exploratory graphs

lobster_traps_clean <- lobster_traps %>% 
  select(YEAR, MONTH, SITE, TRAPS) %>% 
  count(SITE, YEAR) %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE") %>% 
  rename(traps = "n")

lobster_abundance <- lobster_size %>% 
  select(YEAR, MONTH, SITE, COUNT) %>% 
  count(SITE, YEAR) %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE") %>% 
  rename(abundance = "n")


trap_plot <- lobster_traps %>% 
  select(YEAR, MONTH, SITE, TRAPS) %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE") %>% 
  count(SITE, YEAR) %>% 
  ggplot(aes(x = YEAR, y = n)) +
  geom_line(aes(color=SITE)) +
  theme_classic()


trap_plot
  
```


```{r}

#merge cleaned size and abundance data, graph by site
together <- merge(lobster_traps_clean,lobster_abundance, by = c("SITE", "YEAR")) %>% 
  ggplot(aes(x=traps, y=abundance)) +
  geom_col(aes(fill=SITE))+
  facet_wrap(~SITE, scale = "free")

together

size_traps <- full_join(lobster_traps_clean, lobster_abundance)

size_traps_graph <- ggplot(size_traps, aes(YEAR))+
  geom_line(aes(y=traps, color="traps"))+
  geom_line(aes(y=abundance, color="abundance"))+
  facet_wrap(~SITE)

size_traps_graph

```

2. ANOVA

```{r}

# Compare mean lobster sizes (mm) across five sites for lobster observations collected in 2017

lobster_levene <- leveneTest(SIZE ~ SITE, data = lobster_size2)

lobster_levene

# Variances are not equal

# Can still assume equal variance because largest sample variance is <4x the smallest sample variance, so the variances are close enough (lecture 9)

# ANOVA test for parametric data

# H0: Mean lobster size at the five sites are equal
# HA: At least two means differ significantly

lobster_aov <- aov(SIZE ~ SITE, data = lobster_size2)

summary(lobster_aov)

# At least two samples were taken from populations with different means. Which ones are different? All three different from each other? Or something else?

lob_ph<- TukeyHSD(lobster_aov)

lob_ph

lob_ANOVA <- ggplot(lobster_size2, aes(x = SITE, y = SIZE)) +
  geom_boxplot(fill = "gray", color = "blue") +
  xlab("Site") +
  ylab("Size") +
  theme_bw()
 

lob_ANOVA

```



3. Differences in lobster size between MPAs and nonMPAS in 2012 and 2017  

```{r}
#3: Comparing MPAs v non MPAs between 2012 and 2017 


#create data frames separating MPA sites and nonMPA sites, aggregate size average by year 

class(lobster_size2$SIZE)

mpas_2012 <- lobster_size2 %>% 
  filter(SITE == "IVEE" | SITE == "NAPL", YEAR == "2012") %>% 
  mutate(i = row_number()) %>% 
  spread(SITE,SIZE) %>% 
  select(-i)


nonmpas_2012 <- lobster_size2 %>% 
  filter(SITE == "AQUE" | SITE == "CARP" | SITE == "MOHK", YEAR == "2012") %>% 
  mutate(i = row_number()) %>% 
  spread(SITE,SIZE) %>% 
  select(-i)
 
mpas_2017 <- lobster_size2 %>% 
  filter(SITE == "IVEE" | SITE == "NAPL", YEAR == "2017") %>% 
  mutate(i = row_number()) %>% 
  spread(SITE,SIZE) %>% 
  select(-i)

nonmpas_2017 <- lobster_size2 %>% 
  filter(SITE == "AQUE" | SITE == "CARP" | SITE == "MOHK", YEAR == "2017") %>% 
  mutate(i = row_number()) %>% 
  spread(SITE,SIZE) %>% 
  select(-i)

#First Location: IV MPA
#Test variances to determine what kind of t.test we can use

ftest_iv <- var.test(mpas_2012$IVEE,mpas_2017$IVEE)
ftest_iv

#Variances equal, use Student's T Test

ttest_iv <- t.test(mpas_2012$IVEE,mpas_2017$IVEE, var.equal = TRUE)
ttest_iv

#Second Location: Naples MPA
#Test Variances

ftest_naples <- var.test(mpas_2012$NAPL,mpas_2017$NAPL)
ftest_naples

#Use students T test
ttest_naples <- t.test(mpas_2012$NAPL,mpas_2017$NAPL, var.equal = TRUE)
ttest_naples

#Third Location: Carpinteria Non-MPA
#F test

ftest_carp <- var.test(nonmpas_2012$CARP,nonmpas_2017$CARP)
ftest_carp

#Use students T test
ttest_carp <- t.test(nonmpas_2012$CARP,nonmpas_2017$CARP, var.equal = TRUE)
ttest_carp



#Fourth Location: Arroyo Quemado 
#F Test

ftest_aque <- var.test(nonmpas_2012$AQUE,nonmpas_2017$AQUE)
ftest_aque

#Use students T test

ttest_aque <- t.test(nonmpas_2012$AQUE,nonmpas_2017$AQUE, var.equal = TRUE)
ttest_aque



#Fifth Location: Mohawk Reef
#FTest

ftest_mohk <- var.test(nonmpas_2012$MOHK,nonmpas_2017$MOHK)
ftest_mohk

#Use students T Test

ttest_mohk <- t.test(nonmpas_2012$MOHK,nonmpas_2017$MOHK, var.equal = TRUE)
ttest_mohk

#Only IV and Mohawk Reef showed a significant difference in lobster size between 2012 and 2017. 

#make table with site rows, each year, traps 
```




4. Proportions of legal lobsters
```{r}

carapace <- lobster_size %>% 
  filter(SIZE != "-99999") %>% 
  select(SITE, SIZE) %>% 
  mutate(Legality = ifelse(SIZE >= 86, "Legal", "Illegal")) %>% 
  count(SITE, Legality) %>% 
  spread(Legality, n) %>% 
  select(-SITE)

rownames(carapace) <- c("AQUE", "CARP", "IVEE", "MOHK", "NAPL")

```




Proprtions of legal groups and table of them:

```{r}
carapace_prop <- prop.table(as.matrix(carapace), 1)

carapace_prop

kable(carapace_prop, col.names = c("Illegal","Legal")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)
```

Chi-square test

```{r}
carapace_x2 <- chisq.test(carapace)
carapace_x2

# X2 = 39.011, df = 4, p value = 6.931e-08
```
The proportion of observed lobsters that were above the legal minimum differed signficantly across all sites ($\chi^2$(`r carapace_x2$parameter`) = `r round(carapace_x2$statistic, 2)`, p value < 0.001).
