---
title: "Assignment 4"
author: "Madeline Berger, Sandro Lallas, AnnaClaire Marley"
date: "11/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load necessary packages 

library(tidyverse)
library(vcdExtra)
library(kableExtra)
library(RColorBrewer)
library(car)
library(effsize)
library(kableExtra)
library(formattable)

```

Read in the CSVs

```{r, include = FALSE}


lobster_size <- read_csv("lobster_size_abundance.csv")
lobster_traps <- read_csv("lobster_traps.csv")
avg_size_traps <- read_csv("avg_size_traps .csv")
```

The lobster_size file needs to be converted to tidyverse format

```{r, echo=FALSE}

#convert data to tidyverse
lobster_size1 <- as.data.frame(lobster_size)
lobster_size2 <- expand.dft(lobster_size1, freq = "COUNT") %>% 
  select(YEAR, SITE, SIZE)

```

**Figure 1. Amount of trap buoys at five research sites.**

```{r, echo=FALSE, message = FALSE}

#Data wrangling into tables and exploratory graphs

lobster_traps_clean <- lobster_traps %>% 
  select(YEAR, MONTH, SITE, TRAPS) %>% 
  count(SITE, YEAR) %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE") %>% 
  rename(traps = "n")

traps_summary <- lobster_traps %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE") %>%
  select(YEAR, MONTH, SITE, TRAPS) %>%
  group_by(SITE,YEAR) %>% 
  summarize(
    traps = sum(TRAPS)
  )

lobster_abundance <- lobster_size %>% 
  select(YEAR, MONTH, SITE, COUNT) %>% 
  count(SITE, YEAR) %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE") %>% 
  rename(abundance = "n")


trap_plot <- lobster_traps %>% 
  select(YEAR, MONTH, SITE, TRAPS) %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE") %>% 
  count(SITE, YEAR) %>% 
  ggplot(aes(x = YEAR, y = n)) +
  geom_line(aes(color=SITE)) +
  theme_classic() +
  xlab("Year") +
  ylab("Traps") +
  ggtitle("Amount of traps (2012-2017)")



trap_plot
  
```


```{r, echo=FALSE, message = FALSE}

#merge cleaned size and abundance data, graph by site
together <- merge(lobster_traps_clean,lobster_abundance, by = c("SITE", "YEAR")) %>% 
  ggplot(aes(x=traps, y=abundance)) +
  geom_col(aes(fill=SITE))+
  facet_wrap(~SITE, scale = "free") +
  ggtitle("Amount of traps at each site (2012-2017)") +
  labs(
    x = "Year",
    y = "Traps"
  ) +
  scale_y_continuous(expand = c(0,0))

together


# lobster traps across sites
trap_plot <- lobster_traps %>% 
  select(YEAR, MONTH, SITE, TRAPS) %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE") %>% 
  count(SITE, YEAR) %>% 
  ggplot(aes(x = YEAR, y = n)) +
  geom_line(aes(color=SITE)) +
  theme_classic() +
  scale_x_continuous(expand=c(0,0)) +
  ggtitle("Amount of traps at each site (2012-2017)") +
  labs(
    x = "Year",
    y = "Traps"
  )

trap_plot

# lobster abundance across sites
abundance_plot <- lobster_size %>% 
  select(YEAR, MONTH, SITE, COUNT) %>% 
  count(SITE, YEAR) %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE") %>% 
  rename(abundance = "n") %>% 
  ggplot(aes(x = YEAR, y = abundance)) +
  geom_line(aes(color=SITE)) +
  theme_classic() +
  scale_x_continuous(expand=c(0,0)) +
  ggtitle("Abundance of lobsters at each site (2012-2017)") +
  labs(
    x = "Year",
    y = "Lobsters"
  )

abundance_plot


#made a different merge for the graph - this way makes sure no data is lost 
abundance_traps <- full_join(traps_summary, lobster_abundance)

#use new merged data to make side by side line graphs 
abundance_traps_graph <- ggplot(abundance_traps, aes(YEAR))+
  geom_line(aes(y=traps, color="Traps"))+
  geom_line(aes(y=abundance, color="Lobsters"))+
  facet_wrap(~SITE, scale = "free") +
  ggtitle("Amount of traps and lobsters at each site (2012-2017)") +
  labs(
    x = "Year",
    y = "Amount"
  ) +
  scale_y_continuous(expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90))



abundance_traps_graph


#we should write a statement about what this shows 
#aparently we are also supposed to break these out 
```

2. ANOVA

**Figure 3. Lobster size at five research sites (2017).**

```{r, echo = FALSE}

# Compare mean lobster sizes (mm) across five sites for lobster observations collected in 2017

lobster_size3 <- lobster_size2 %>% 
  filter(YEAR == 2017)

lobster_levene <- leveneTest(SIZE ~ SITE, data = lobster_size3)

# Variances are not equal

# Can still assume equal variance because largest sample variance is <4x the smallest sample variance, so the variances are close enough (lecture 9)

# ANOVA test for parametric data

# H0: Mean lobster size at the five sites are equal
# HA: At least two means differ significantly

lobster_aov <- aov(SIZE ~ SITE, data = lobster_size3)

# At least two samples were taken from populations with different means. Which ones are different? All three different from each other? Or something else?

lob_ph<- TukeyHSD(lobster_aov)

# Boxplot graph ANOVA

lob_ANOVA <- ggplot(lobster_size3, aes(x = SITE, y = SIZE)) +
  geom_boxplot(fill = "gray") +
  theme_bw() +
  ggtitle("Mean Lobster Size (mm)") +
  annotate("text", x = 1, y = 110, label = "a,b", family = "Times New Roman") +
  annotate("text", x = 2, y = 125, label = "a", family = "Times New Roman") +
  annotate("text", x = 3, y = 156, label = "a", family = "Times New Roman") +
  annotate("text", x = 4, y = 100, label = "a,b", family = "Times New Roman") +
  annotate("text", x = 5, y = 115, label = "b", family = "Times New Roman") +
  xlab("Site") +
  ylab("Mean lobster size (mm)") +
  scale_y_continuous(lim = c(0, 160), expand = c(0,0)) +
  scale_x_discrete(labels = c("AQUE" = "AQUE\nn=67",
                              "CARP" = "CARP\nn=705",
                              "IVEE" = "IVEE\nn=606",
                              "MOHK" = "MOHK\nn=178",
                              "NAPL" = "NAPL\nn=112"
                                ))
 
lob_ANOVA

# Data preparation for column graph ANOVA 

lob_col_data <- lobster_size3 %>%
  group_by(SITE) %>%
  summarize(
    mean = mean(SIZE),
    sd = sd(SIZE),
    length = length(SIZE)
  )

# Column graph ANOVA

lob_col <- ggplot(lob_col_data, aes(x = SITE, y = mean)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean+sd), width = .2) +
  theme_bw() +
  ggtitle("Mean Lobster Size (mm)") +
  annotate("text", x = 1, y = 90, label = "a,b", family = "Times New Roman") +
  annotate("text", x = 2, y = 89, label = "a", family = "Times New Roman") +
  annotate("text", x = 3, y = 89, label = "a", family = "Times New Roman") +
  annotate("text", x = 4, y = 85, label = "a,b", family = "Times New Roman") +
  annotate("text", x = 5, y = 91, label = "b", family = "Times New Roman") +
  xlab("Site") +
  ylab("Mean lobster size (mm)") +
  scale_y_continuous(lim = c(0, 100), expand = c(0,0)) +
  scale_x_discrete(labels = c("AQUE" = "AQUE\nn=67",
                              "CARP" = "CARP\nn=705",
                              "IVEE" = "IVEE\nn=606",
                              "MOHK" = "MOHK\nn=178",
                              "NAPL" = "NAPL\nn=112"
                                ))

```


Figure 2. In 2017, mean lobster carapace size (mm) differed significantly across the five sampled sites (by one-way ANOVA with Tukey's HSD, F(4, 1663) = 3.424, *p* < 0.001, with $\alpha$ = 0.05 throughout). Post-hoc analysis by Tukey's HSD revealed that mean lobster size at Naples Reef differed significantly from Carpinteria and Isla Vista (pairwise *p* < 0.001) and had the highest mean lobster size (76.23 mm). Meanwhile, mean lobster size at Arroyo Quemado and Mohawk were not significantly different from any of the other sites (pairwise *p* > 0.05).


3. Differences in lobster size between MPAs and nonMPAS in 2012 and 2017  

```{r, echo = FALSE}
#3: Comparing MPAs v non MPAs between 2012 and 2017 


#create data frames separating MPA sites and nonMPA sites

class(lobster_size2$SIZE)

mpas_2012 <- lobster_size2 %>% 
  filter(SITE == "IVEE" | SITE == "NAPL", YEAR == "2012") %>% 
  mutate(i = row_number()) %>% 
  spread(SITE,SIZE) %>% 
  select(-i)


nonmpas_2012 <- lobster_size2 %>% 
  filter(SITE == "AQUE" | SITE == "CARP" | SITE == "MOHK", YEAR == "2012") %>% 
  mutate(i = row_number()) %>% 
  spread(SITE,SIZE) %>% 
  select(-i)
 
mpas_2017 <- lobster_size2 %>% 
  filter(SITE == "IVEE" | SITE == "NAPL", YEAR == "2017") %>% 
  mutate(i = row_number()) %>% 
  spread(SITE,SIZE) %>% 
  select(-i)

nonmpas_2017 <- lobster_size2 %>% 
  filter(SITE == "AQUE" | SITE == "CARP" | SITE == "MOHK", YEAR == "2017") %>% 
  mutate(i = row_number()) %>% 
  spread(SITE,SIZE) %>% 
  select(-i)


sample_sizes2017 <- lobster_size %>% 
  filter(SITE == "AQUE" | SITE == "CARP" | SITE == "MOHK" | SITE =="IVEE"| SITE == "NAPL", YEAR == "2017") %>% 
  group_by(SITE) %>% 
  summarize(
    sizes2017 = length(SIZE)
  )

sample_sizes2017

sample_sizes2012 <- lobster_size %>% 
  filter(SITE == "AQUE" | SITE == "CARP" | SITE == "MOHK" | SITE =="IVEE"| SITE == "NAPL", YEAR == "2012") %>% 
  group_by(SITE) %>% 
  summarize(
    sizes2012 = length(SIZE)
  )

sample_sizes2012
```




```{r, include = FALSE}
#create exploratory graphs 

#MPAs, 2012 
hist_mpa2012 <- lobster_size2 %>% 
  filter(SITE == "IVEE" | SITE == "NAPL", YEAR == "2012") %>% 
  ggplot(aes(x=SIZE))+
  geom_histogram(bins = 5)+
  facet_wrap(~SITE)

hist_mpa2012
 

qq_mpa2012 <- lobster_size2 %>% 
  filter(SITE == "IVEE" | SITE == "NAPL", YEAR == "2012") %>% 
  ggplot(aes(sample=SIZE))+
  geom_qq()+
  facet_wrap(~SITE)

qq_mpa2012

#IV does not look normally distributed in 2012
#MPAS 2017

hist_mpa2017 <- lobster_size2 %>% 
  filter(SITE == "IVEE" | SITE == "NAPL", YEAR == "2017") %>% 
  ggplot(aes(x=SIZE))+
  geom_histogram(bins = 5)+
  facet_wrap(~SITE)

hist_mpa2017
 

qq_mpa2017 <- lobster_size2 %>% 
  filter(SITE == "IVEE" | SITE == "NAPL", YEAR == "2017") %>% 
  ggplot(aes(sample=SIZE))+
  geom_qq()+
  facet_wrap(~SITE)

qq_mpa2017


#nonMPAS, 2012

hist_nonmpa2012 <- lobster_size2 %>% 
  filter(SITE == "AQUE" | SITE == "MOHK" | SITE == "CARP", YEAR == "2012") %>% 
  ggplot(aes(x=SIZE))+
  geom_histogram(bins = 10)+
  facet_wrap(~SITE)

hist_nonmpa2012


qq_nonmpa2012 <- lobster_size2 %>% 
  filter(SITE == "AQUE" | SITE == "MOHK" | SITE == "CARP", YEAR == "2012") %>% 
  ggplot(aes(sample=SIZE))+
  geom_qq()+
  facet_wrap(~SITE)
qq_nonmpa2012

#nonMPAS2017


hist_nonmpa2017 <- lobster_size2 %>% 
  filter(SITE == "AQUE" | SITE == "MOHK" | SITE == "CARP", YEAR == "2017") %>% 
  ggplot(aes(x=SIZE))+
  geom_histogram(bins = 10)+
  facet_wrap(~SITE)

hist_nonmpa2017


qq_nonmpa2017 <- lobster_size2 %>% 
  filter(SITE == "AQUE" | SITE == "MOHK" | SITE == "CARP", YEAR == "2017") %>% 
  ggplot(aes(sample=SIZE))+
  geom_qq()+
  facet_wrap(~SITE)

qq_nonmpa2017

```

```{r, echo = FALSE}
#Make 10 vectors to use for effectsize

#IV
iv2012 <- lobster_size2 %>% 
  filter(SITE == "IVEE", YEAR == "2012") %>% 
  pull(SIZE)

iv2017 <- lobster_size2 %>% 
  filter(SITE == "IVEE", YEAR == "2017") %>% 
  pull(SIZE)

#Naples

napl2012 <- lobster_size2 %>% 
  filter(SITE == "NAPL", YEAR == "2012") %>% 
  pull(SIZE)

napl2017 <- lobster_size2 %>% 
  filter(SITE == "NAPL", YEAR == "2017") %>% 
  pull(SIZE)

#Carpinteria 

carp2012 <- lobster_size2 %>% 
  filter(SITE == "CARP", YEAR == "2012") %>% 
  pull(SIZE)

carp2017 <- lobster_size2 %>% 
  filter(SITE == "CARP", YEAR == "2017") %>% 
  pull(SIZE)

#Arroyo Quemado

aque2012 <- lobster_size2 %>% 
  filter(SITE == "AQUE", YEAR == "2012") %>% 
  pull(SIZE)

aque2017 <- lobster_size2 %>% 
  filter(SITE == "AQUE", YEAR == "2017") %>% 
  pull(SIZE)



#Mohawk

mohk2012 <- lobster_size2 %>% 
  filter(SITE == "MOHK", YEAR == "2012") %>% 
  pull(SIZE)

mohk2017 <- lobster_size2 %>% 
  filter(SITE == "MOHK", YEAR == "2017") %>% 
  pull(SIZE)
```

```{r, include = FALSE}
#create values of mean and sd

#IV
mean_iv12<- mean(iv2012)
mean_iv12

sd_iv12 <- sd(iv2012)
sd_iv12


mean_iv17<- mean(iv2017)
mean_iv17

sd_iv17 <- sd(iv2017)
sd_iv17

#Naples

mean_napl12<- mean(napl2012)
mean_napl12

sd_napl12 <- sd(napl2012)
sd_napl12


mean_napl17<- mean(napl2017)
mean_napl17

sd_napl17 <- sd(napl2017)
sd_napl17


#Arroyo Quemado

mean_aque12<- mean(aque2012)
mean_aque12

sd_aque12 <- sd(aque2012)
sd_aque12


mean_aque17<- mean(aque2017)
mean_aque17

sd_aque17 <- sd(aque2017)
sd_aque17


#Carpinteria

mean_carp12<- mean(carp2012)
mean_carp12

sd_carp12 <- sd(carp2012)
sd_carp12


mean_carp17<- mean(carp2017)
mean_carp17

sd_carp17 <- sd(carp2017)
sd_carp17

#Mohawk

mean_mohk12<- mean(mohk2012)
mean_mohk12

sd_mohk12 <- sd(mohk2012)
sd_mohk12


mean_mohk17<- mean(mohk2017)
mean_mohk17

sd_mohk17 <- sd(mohk2017)
sd_mohk17
```



```{r, echo = FALSE, results = FALSE}
#First Location: IV MPA
#Test variances to determine what kind of t.test we can use

ftest_iv <- var.test(mpas_2012$IVEE,mpas_2017$IVEE)
ftest_iv

#Variances equal, use Student's T Test

ttest_iv <- t.test(mpas_2012$IVEE,mpas_2017$IVEE, var.equal = TRUE)
ttest_iv

#Siglevel = 0.05, no significant difference 

#calculate effect size using Cohen's D

effsize_iv <- cohen.d(iv2012,iv2017)
effsize_iv


#Experiment: try Mann Whitney U since data did not appear normal

manu_iv <- wilcox.test(mpas_2012$IVEE,mpas_2017$IVEE)
manu_iv
```

Average lobster size at Isla Vista Reef in 2012 (n = 26) and 2017 (n = 606) did not differed significantly (t{630} = `r {round(ttest_iv$statistic,2)}`, p = `r {round(ttest_iv$p.value,3)}`, $\alpha$ = 0.05). To futher examine the difference between the two years, we calculated the effect size using Cohen's D. The result, `r {round(effsize_iv$estimate,2)}`, indicates a small effect size, and the aboslute difference in mean size between the two years was??. 



```{r, echo=FALSE, results = FALSE}
#Second Location: Naples MPA
#Test Variances

ftest_naples <- var.test(mpas_2012$NAPL,mpas_2017$NAPL)
ftest_naples

#Use students T test
ttest_naples <- t.test(mpas_2012$NAPL,mpas_2017$NAPL, var.equal = TRUE)
ttest_naples

#No significant difference
#Cohen's d for effect size
effsize_naples <- cohen.d(napl2012,napl2017)
effsize_naples

```



Average lobster size at Naples Reef MPA in 2012 (n = 6) and 2017 (n = 112) did not differed significantly (t{5} = `r {round(ttest_naples$statistic,2)}`, p = `r {round(ttest_naples$p.value,3)}`, $\alpha$ = 0.05). To futher examine the difference between the two years, we calculated the effect size using Cohen's D. The result, `r {round(effsize_naples$estimate,2)}`, indicates a small effect size, and the aboslute difference in mean size between the two years was....


```{r, echo=FALSE, results = FALSE}
#Third Location: Carpinteria Non-MPA
#F test

ftest_carp <- var.test(nonmpas_2012$CARP,nonmpas_2017$CARP)
ftest_carp

#Use students T test
ttest_carp <- t.test(nonmpas_2012$CARP,nonmpas_2017$CARP, var.equal = TRUE)
ttest_carp


#Cohen's d for effect size
effsize_carp <- cohen.d(carp2012,carp2017)
effsize_carp


```

Average lobster size at the Carpinteria site in 2012 (n = 78) and 2017 (n = 704 ) did not differ significantly (t{5} = `r {round(ttest_carp$statistic,2)}`, p = `r {round(ttest_carp$p.value,3)}`, $\alpha$ = 0.05). To futher examine the difference between the two years, we calculated the effect size using Cohen's D. The result, `r {round(effsize_carp$estimate,2)}`, indicates an extremely small effect size, and the aboslute difference in mean size between the two years was....




```{r, echo=FALSE, results = FALSE}
#Fourth Location: Arroyo Quemado 
#F Test

ftest_aque <- var.test(nonmpas_2012$AQUE,nonmpas_2017$AQUE)
ftest_aque

#Use students T test

ttest_aque <- t.test(nonmpas_2012$AQUE,nonmpas_2017$AQUE, var.equal = TRUE)
ttest_aque

#Cohen's d for effect size
effsize_aque <- cohen.d(aque2012,aque2017)
effsize_aque

#Mann Whitney U
manu_aque <- wilcox.test(nonmpas_2012$AQUE,nonmpas_2017$AQUE)
manu_aque

```


Average lobster size at Arroyo Quemado Reef in 2012 (n = 38) and 2017 (n = 67 ) did not differ significantly (t{5} = `r {round(ttest_aque$statistic,2)}`, p = `r {round(ttest_aque$p.value,3)}`, $\alpha$ = 0.05). To futher examine the difference between the two years, we calculated the effect size using Cohen's D. The result, `r {round(effsize_aque$estimate,2)}`, indicates a small effect size, and the aboslute difference in mean size between the two years was....


```{r, include = FALSE}
#Fifth Location: Mohawk Reef
#FTest

ftest_mohk <- var.test(nonmpas_2012$MOHK,nonmpas_2017$MOHK)
ftest_mohk

#Use students T Test

ttest_mohk <- t.test(nonmpas_2012$MOHK,nonmpas_2017$MOHK, var.equal = TRUE)
ttest_mohk


#Cohen's d for effect size
effsize_mohk <- cohen.d(mohk2012,mohk2017)
effsize_mohk


#make table with site rows, each year, traps 
```

Average lobster size at Mohawk Reef in 2012 (n = 82) and 2017 (n = 177) did differ significantly (t{5} = `r {round(ttest_mohk$statistic,2)}`, p = `r {round(ttest_mohk$p.value,3)}`, $\alpha$ = 0.05). To futher examine the difference between the two years, we calculated the effect size using Cohen's D. The result, `r {round(effsize_mohk$estimate,2)}`, indicates a medium effect size, and the aboslute difference in mean size between the two years was....



```{r, include = FALSE}
#Summary Table of  average sizes
size_summary_table <- lobster_size2 %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE", YEAR == "2012" | YEAR == "2017") %>%
  group_by(SITE,YEAR) %>% 
  summarize(
    avg = mean(SIZE)
  ) %>% 
  mutate(i = row_number()) %>% 
  spread(YEAR,avg) %>% 
  select(-i)

  
  
size_summary_table


#Summary table of traps per year 
  
traps_summary_table <- traps_summary %>% 
  filter(SITE== "CARP" | SITE== "MOHK" | SITE== "AQUE" | SITE== "NAPL" | SITE== "IVEE", YEAR == "2012" | YEAR == "2017") %>%
  group_by(SITE, YEAR) %>% 
  summarize(
    avg = mean(traps)
  ) %>% 
  mutate(i = row_number()) %>% 
  spread(YEAR,avg) %>% 
  select(-i)

traps_summary_table

```

```{r, echo = FALSE, message = FALSE}


size_traps_summary <- avg_size_traps %>% 
  mutate(size_difference = round(size_2017-size_2012,2)) %>% 
  mutate(traps_difference = traps_2017-traps_2012) %>% 
  select(-Site) %>% 
  select(-traps_2017, -traps_2012) %>% 
  rename("Size in 2012" = size_2012, "Size in 2017" = size_2017, "Change in Size" = size_difference, "Change in Traps" = traps_difference)



rownames(size_traps_summary) <- c("Isla Vista","Naples Reef", "Arroyo Quemado", "Carpinteria","Mohawk Reef")


formattable(size_traps_summary, list(`Change in Size`= formatter("span", style = x ~ style(color = ifelse(x > 0, "green", ifelse(x < 0, "red", "black"))))))




```



4. Proportions of legal lobsters
```{r, echo=FALSE, message = FALSE}
#New Pretty Table for display in report
sample_number <- c(67, 705, 606, 178, 112)

carapace <- lobster_size2 %>% 
  filter(YEAR == "2017") %>% 
  select(SITE, SIZE) %>% 
  mutate(Legality = ifelse(SIZE >= 86, "Legal", "Illegal")) %>% 
  count(SITE, Legality) %>% 
  spread(Legality, n) %>% 
  select(-SITE) %>% 
  mutate(Sample_Size = sample_number) %>% 
  mutate(Prop_legal = round(Legal/Sample_Size, 2)) %>% 
  select(-Sample_Size)
  
rownames(carapace) <- c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples")

carapace_table2 <- kable(carapace, col.names = c("Illegal Size","Legal Size", "Proportion Legal"), align = "c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)

carapace_table2 


carapace2 <- lobster_size2 %>% 
  filter(YEAR == "2017") %>% 
  select(SITE, SIZE) %>% 
  mutate(Legality = ifelse(SIZE >= 86, "Legal", "Illegal")) %>% 
  count(SITE, Legality) %>% 
  spread(Legality, n) %>% 
  select(-SITE)
  
rownames(carapace2) <- c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk", "Naples")

```



Proprtions of legal groups and table of them:

```{r, echo=FALSE}
carapace_prop <- prop.table(as.matrix(carapace), 1) %>% 
  round(2)

carapace_prop

carapace_prop_table <- kable(carapace_prop, col.names = c("Illegal Size","Legal Size", "Sample Size"), align = "c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)

carapace_prop_table

```

Chi-square test

<<<<<<< HEAD
```{r, echo=FALSE}
carapace_x2 <- chisq.test(carapace2)
carapace_x2


```
The proportion of observed lobsters that were above the legal minimum differed signficantly across all sites ($\chi^2$(`r carapace_x2$parameter`) = `r round(carapace_x2$statistic, 2)`, p value < 0.001).
